---
title: "Longitudinal Difference-in-Differences Syllable Usage Analysis"
author: "MoSeq Analysis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
params:
  data_dir: "data/final_model_kappa_40000000"
  counts_csv: "data/syllable_counts_did.csv"
  rebuild_counts: false
  python_bin: "/opt/miniconda3/envs/moseq2-app/bin/python"
  labels_xlsx: "data/Mo-seq syllable behavior labeling.xlsx"
  labels_sheet: "Copy of Kaitlyn + Steve (KH)"
  baseline_timepoint: "b2"
  post_timepoints: ["wk1", "wk2", "wk3", "washout"]
  max_syllable_id: 54
  fdr_method: "BH"
  min_total_counts: 1
  # Drop visually unstable contrasts from figures (still kept in CSV)
  max_abs_effect_plot: 8
  max_se_plot: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 5
)
```

# Objective

Fit a **syllable-specific beta-binomial GLMM** to mouse-level syllable usage and estimate:

1. Sex-specific difference-in-differences (DiD) for CORT vs VEH relative to baseline
2. The three-way difference (DDD) testing whether DiD differs by sex

This implements the longitudinal DiD / DDD formulation with mouse random intercepts.

# Model

For each syllable $s$, mouse $i$, and timepoint $t$:

$$
Y_{ist} \sim \mathrm{BetaBinomial}(N_{it}, p_{ist}, \phi_s)
$$

$$
\begin{aligned}
\mathrm{logit}(p_{ist})
&= \beta_{0s}
+ \beta_{Ts} T_i
+ \beta_{Ss} S_i
+ \sum_{j=1}^{J} \beta_{js} I_{jt} \\
&\quad
+ \sum_{j=1}^{J} \beta_{Tj,s}(T_i I_{jt})
+ \beta_{TS,s}(T_i S_i)
+ \sum_{j=1}^{J} \beta_{jS,s}(I_{jt} S_i) \\
&\quad
+ \sum_{j=1}^{J} \beta_{TjS,s}(T_i I_{jt} S_i)
+ b_{is}
\end{aligned}
$$

$$
b_{is} \sim \mathcal{N}(0, \sigma_{b,s}^2)
$$

Coding:

- $Y_{ist}$: onset count of syllable $s$ for mouse $i$ at time $t$
- $N_{it}$: total syllable onsets for that session
- $T_i = 1$ for CORT, $0$ for VEH
- $S_i = 1$ for male, $0$ for female
- $I_{jt}$: indicators for post-baseline timepoints; baseline (`r params$baseline_timepoint`) is the reference

Contrasts (female = reference sex):

$$
\mathrm{DiD}_{\mathrm{Female},j}^{(s)} = \beta_{Tj,s}
$$

$$
\mathrm{DiD}_{\mathrm{Male},j}^{(s)} = \beta_{Tj,s} + \beta_{TjS,s}
$$

$$
\mathrm{DDD}_{j}^{(s)} = \beta_{TjS,s}
$$

# Setup

```{r packages}
library(tidyverse)
library(glmmTMB)
library(readxl)
```

```{r paths}
repo_root <- if (requireNamespace("rprojroot", quietly = TRUE)) {
  tryCatch(
    rprojroot::find_root(rprojroot::is_git_root),
    error = function(e) normalizePath(".", winslash = "/", mustWork = TRUE)
  )
} else {
  normalizePath(".", winslash = "/", mustWork = TRUE)
}

data_dir <- if (grepl("^(/|[A-Za-z]:)", params$data_dir)) {
  params$data_dir
} else {
  file.path(repo_root, params$data_dir)
}
counts_csv <- if (grepl("^(/|[A-Za-z]:)", params$counts_csv)) {
  params$counts_csv
} else {
  file.path(repo_root, params$counts_csv)
}
labels_xlsx <- if (grepl("^(/|[A-Za-z]:)", params$labels_xlsx)) {
  params$labels_xlsx
} else {
  file.path(repo_root, params$labels_xlsx)
}
if (!file.exists(labels_xlsx)) {
  stop("Syllable labels file not found: ", labels_xlsx)
}
if (!dir.exists(data_dir)) {
  stop("data_dir not found: ", data_dir)
}
prep_script <- file.path(repo_root, "scripts", "prepare_syllable_counts_did.py")

baseline <- params$baseline_timepoint
post_tps <- unlist(params$post_timepoints)
time_levels <- c(baseline, post_tps)

cat("data_dir: ", data_dir, "\n", sep = "")
cat("counts_csv: ", counts_csv, "\n", sep = "")
cat("labels_xlsx: ", labels_xlsx, "\n", sep = "")
cat("timepoints: ", paste(time_levels, collapse = ", "), "\n", sep = "")
```

# 1. Data preparation from `moseq_df.csv`

`moseq_df.csv` is frame-level MoSeq output (wiki: one row per frame). For this model we need session-level counts:

- Keep frames with `onset == TRUE` and `labels (usage sort) >= 0` (drop the AR-HMM init label `-5`)
- $Y_{ist}$: number of onsets of syllable $s$ in the session
- $N_{it}$: total onsets in the session
- Mouse ID: `SubjectName`, with batch-labeled names recovered from `SessionName`
- Design factors parsed from `group` names of the form `{treatment}_{timepoint}_{sex}`

Exclude groups containing `ket` or `sal` so the comparison is CORT vs VEH over the longitudinal time course.

```{r rebuild-counts}
if (isTRUE(params$rebuild_counts) || !file.exists(counts_csv)) {
  if (!file.exists(prep_script)) {
    stop("Missing prep script: ", prep_script)
  }
  dir.create(dirname(counts_csv), recursive = TRUE, showWarnings = FALSE)
  cmd <- sprintf(
    '"%s" "%s" --data-dir "%s" --out "%s" --timepoints %s',
    params$python_bin,
    prep_script,
    data_dir,
    counts_csv,
    paste(time_levels, collapse = " ")
  )
  cat("Running:\n", cmd, "\n", sep = "")
  status <- system(cmd)
  if (!identical(as.integer(status), 0L)) {
    stop("Count preparation failed with status ", status)
  }
} else {
  cat("Using existing counts file.\n")
}
```

# 2. Load analysis table and check design

```{r load-counts}
raw <- read_csv(counts_csv, show_col_types = FALSE)

did_df <- raw %>%
  mutate(
    treatment = factor(treatment, levels = c("veh", "cort")),
    sex = factor(sex, levels = c("f", "m")),
    timepoint = factor(timepoint, levels = time_levels),
    mouse_id = factor(mouse_id),
    syllable = as.integer(syllable)
  ) %>%
  filter(
    !is.na(timepoint),
    N >= params$min_total_counts,
    is.na(params$max_syllable_id) | syllable <= params$max_syllable_id
  )

dup_mouse_time <- did_df %>%
  distinct(mouse_id, timepoint, uuid) %>%
  count(mouse_id, timepoint) %>%
  filter(n > 1)
if (nrow(dup_mouse_time) > 0) {
  stop("Duplicate mouse_id x timepoint sessions present")
}

session_n <- did_df %>%
  distinct(uuid, mouse_id, treatment, sex, timepoint, group, N)

cat("Sessions:", nrow(session_n), "\n")
cat("Mice:", n_distinct(session_n$mouse_id), "\n")
cat("Syllables:", n_distinct(did_df$syllable), "\n")
```

```{r design-table}
session_n %>%
  count(treatment, sex, timepoint, name = "n_sessions") %>%
  pivot_wider(names_from = timepoint, values_from = n_sessions, values_fill = 0) %>%
  knitr::kable(caption = "Sessions per treatment × sex × timepoint")
```

```{r completeness}
session_n %>%
  count(mouse_id, name = "n_timepoints") %>%
  count(n_timepoints, name = "n_mice") %>%
  knitr::kable(caption = "Longitudinal completeness (timepoints observed per mouse)")
```

Usage reconstructed from onset counts should match MoSeq `stats_df` usage (same definition). Optional check if `stats_df.csv` is present:

```{r usage-check}
stats_path <- file.path(data_dir, "stats_df.csv")
if (file.exists(stats_path)) {
  stats_df <- read_csv(stats_path, show_col_types = FALSE) %>%
    select(uuid, syllable, usage_stats = usage)
  cmp <- did_df %>%
    select(uuid, syllable, usage_counts = usage) %>%
    inner_join(stats_df, by = c("uuid", "syllable")) %>%
    mutate(abs_diff = abs(usage_counts - usage_stats))
  cat(
    "Median |usage_counts - usage_stats| = ",
    format(median(cmp$abs_diff), digits = 3),
    "; max = ",
    format(max(cmp$abs_diff), digits = 3),
    "\n",
    sep = ""
  )
} else {
  cat("stats_df.csv not found; skipping usage cross-check.\n")
}
```

# 3. Fit syllable-specific beta-binomial GLMMs

Model formula (R):

```r
cbind(y, N - y) ~ treatment * timepoint * sex + (1 | mouse_id)
```

with `family = betabinomial()`. Female / VEH / baseline are reference levels, so coefficient names map directly to the DiD / DDD contrasts above.

```{r fit-helpers}
extract_effects <- function(fit, syllable_id, post_tps) {
  b <- fixef(fit)$cond
  V <- vcov(fit)$cond
  se <- sqrt(diag(V))

  rows <- lapply(post_tps, function(tp) {
    fem_nm <- paste0("treatmentcort:timepoint", tp)
    ddd_nm <- paste0("treatmentcort:timepoint", tp, ":sexm")
    if (!all(c(fem_nm, ddd_nm) %in% names(b))) {
      return(NULL)
    }

    did_f <- unname(b[fem_nm])
    did_f_se <- unname(se[fem_nm])
    ddd <- unname(b[ddd_nm])
    ddd_se <- unname(se[ddd_nm])
    did_m <- did_f + ddd
    did_m_se <- sqrt(did_f_se^2 + ddd_se^2 + 2 * V[fem_nm, ddd_nm])

    sigma_b <- tryCatch(
      unname(attr(VarCorr(fit)$cond$mouse_id, "stddev")),
      error = function(e) NA_real_
    )

    tibble(
      syllable = syllable_id,
      timepoint = tp,
      DiD_Female = did_f,
      DiD_Female_SE = did_f_se,
      DiD_Female_z = did_f / did_f_se,
      DiD_Female_p = 2 * pnorm(-abs(did_f / did_f_se)),
      DiD_Male = did_m,
      DiD_Male_SE = did_m_se,
      DiD_Male_z = did_m / did_m_se,
      DiD_Male_p = 2 * pnorm(-abs(did_m / did_m_se)),
      DDD = ddd,
      DDD_SE = ddd_se,
      DDD_z = ddd / ddd_se,
      DDD_p = 2 * pnorm(-abs(ddd / ddd_se)),
      sigma_b = sigma_b,
      AIC = AIC(fit),
      converged = isTRUE(fit$fit$convergence == 0)
    )
  })
  bind_rows(rows)
}

fit_syllable <- function(dat, syllable_id, post_tps) {
  sdat <- dat %>% filter(syllable == syllable_id)
  # Skip syllables never observed
  if (sum(sdat$y) == 0) {
    return(list(effects = NULL, error = "all-zero counts"))
  }
  fit <- tryCatch(
    glmmTMB(
      cbind(y, N - y) ~ treatment * timepoint * sex + (1 | mouse_id),
      data = sdat,
      family = betabinomial()
    ),
    error = function(e) e
  )
  if (inherits(fit, "error")) {
    return(list(effects = NULL, error = conditionMessage(fit)))
  }
  list(effects = extract_effects(fit, syllable_id, post_tps), error = NULL, fit = fit)
}
```

```{r fit-all, cache=TRUE}
syllables <- sort(unique(did_df$syllable))
fit_log <- tibble(syllable = integer(), status = character(), detail = character())
effects_list <- list()

for (s in syllables) {
  res <- fit_syllable(did_df, s, post_tps)
  if (is.null(res$effects)) {
    fit_log <- add_row(
      fit_log,
      syllable = s,
      status = "failed",
      detail = if (is.null(res$error)) "unknown" else res$error
    )
  } else {
    effects_list[[as.character(s)]] <- res$effects
    fit_log <- add_row(
      fit_log,
      syllable = s,
      status = if (all(res$effects$converged)) "ok" else "converged_flag",
      detail = NA_character_
    )
  }
}

effects <- bind_rows(effects_list)
cat("Fitted syllables:", n_distinct(effects$syllable), "/", length(syllables), "\n")
fit_log %>% count(status) %>% knitr::kable(caption = "Model fit status by syllable")
```

# 4. Multiple-testing adjustment

FDR is applied separately within each contrast family across syllables × post-baseline timepoints.
Rows with non-finite standard errors (typically singular / non-converged fits) are retained in the
output file but excluded from FDR adjustment and from the primary significance tables.

```{r fdr}
effects <- effects %>%
  mutate(
    estimable = is.finite(DiD_Female_SE) & is.finite(DiD_Male_SE) & is.finite(DDD_SE) &
      converged %in% TRUE,
    DiD_Female_q = NA_real_,
    DiD_Male_q = NA_real_,
    DDD_q = NA_real_
  )

idx <- which(effects$estimable)
effects$DiD_Female_q[idx] <- p.adjust(effects$DiD_Female_p[idx], method = params$fdr_method)
effects$DiD_Male_q[idx] <- p.adjust(effects$DiD_Male_p[idx], method = params$fdr_method)
effects$DDD_q[idx] <- p.adjust(effects$DDD_p[idx], method = params$fdr_method)

out_dir <- file.path(repo_root, "figures")
eps_dir <- file.path(out_dir, "eps")
png_dir <- file.path(out_dir, "png")
dir.create(eps_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(png_dir, showWarnings = FALSE, recursive = TRUE)

save_plot_eps_png <- function(plot, stem, width, height, dpi = 300) {
  stem <- sub("\\.(eps|png|pdf)$", "", basename(stem), ignore.case = TRUE)
  eps_path <- file.path(eps_dir, paste0(stem, ".eps"))
  png_path <- file.path(png_dir, paste0(stem, ".png"))
  ggsave(eps_path, plot, width = width, height = height, device = "eps")
  ggsave(png_path, plot, width = width, height = height, dpi = dpi)
  cat("Saved: ", eps_path, "\n", sep = "")
  cat("Saved: ", png_path, "\n", sep = "")
  invisible(c(eps = eps_path, png = png_path))
}

effects_path <- file.path(repo_root, "data", "did_ddd_effects.csv")
write_csv(effects, effects_path)
cat("Wrote ", effects_path, "\n", sep = "")
cat("Estimable contrasts: ", sum(effects$estimable), "/", nrow(effects), "\n", sep = "")
cat("Figures (EPS): ", eps_dir, "\n", sep = "")
cat("Figures (PNG): ", png_dir, "\n", sep = "")
```

# 5. Results

Effects are on the **logit (log-odds) scale**. Positive DiD means CORT increased syllable usage relative to VEH more than at baseline (for that sex). Positive DDD means the male DiD exceeds the female DiD.

For figures, keep `estimable` contrasts and further drop extreme / high-SE estimates that distort the axis. Exact values stay in `did_ddd_effects.csv` and the tables below.

```{r load-categories}
cat_levels <- c("locomotion", "exploratory", "grooming", "immobility", "hesitation")

syllable_cats <- readxl::read_excel(
  labels_xlsx,
  sheet = params$labels_sheet,
  skip = 1
) %>%
  transmute(
    syllable = as.integer(`syllable short id`),
    category = str_to_lower(str_trim(`general category`))
  ) %>%
  filter(!is.na(syllable), category %in% cat_levels) %>%
  distinct(syllable, .keep_all = TRUE)

effects <- effects %>%
  left_join(syllable_cats, by = "syllable") %>%
  mutate(
    category = factor(category, levels = cat_levels),
    timepoint = factor(timepoint, levels = post_tps)
  )

plot_df <- effects %>%
  filter(
    estimable,
    !is.na(category),
    abs(DDD) <= params$max_abs_effect_plot,
    abs(DiD_Female) <= params$max_abs_effect_plot,
    abs(DiD_Male) <= params$max_abs_effect_plot,
    DDD_SE <= params$max_se_plot,
    DiD_Female_SE <= params$max_se_plot,
    DiD_Male_SE <= params$max_se_plot
  )

cat(
  "Plotting ", nrow(plot_df), " / ", sum(effects$estimable),
  " estimable contrasts after stability filters.\n",
  sep = ""
)
```

```{r summary-counts}
effects %>%
  filter(estimable) %>%
  summarise(
    n_tests = n(),
    DiD_Female_p05 = sum(DiD_Female_p < 0.05, na.rm = TRUE),
    DiD_Female_q05 = sum(DiD_Female_q < 0.05, na.rm = TRUE),
    DiD_Male_p05 = sum(DiD_Male_p < 0.05, na.rm = TRUE),
    DiD_Male_q05 = sum(DiD_Male_q < 0.05, na.rm = TRUE),
    DDD_p05 = sum(DDD_p < 0.05, na.rm = TRUE),
    DDD_q05 = sum(DDD_q < 0.05, na.rm = TRUE)
  ) %>%
  knitr::kable(caption = "Significant estimable tests at raw p < 0.05 and FDR q < 0.05")
```

## Main figure: sex-specific DiD by category

Show **both sexes**. DDD alone does not say whether CORT changed usage in females, males, or both.

```{r did-category-plot, fig.width=12, fig.height=8}
did_long <- plot_df %>%
  select(
    syllable, category, timepoint,
    Female = DiD_Female, Male = DiD_Male,
    Female_SE = DiD_Female_SE, Male_SE = DiD_Male_SE,
    Female_q = DiD_Female_q, Male_q = DiD_Male_q
  ) %>%
  pivot_longer(
    cols = c(Female, Male),
    names_to = "sex",
    values_to = "DiD"
  ) %>%
  mutate(
    SE = if_else(sex == "Female", Female_SE, Male_SE),
    q = if_else(sex == "Female", Female_q, Male_q),
    syllable = factor(syllable, levels = sort(unique(syllable)))
  )

p_did <- ggplot(
  did_long,
  aes(
    x = syllable,
    y = DiD,
    ymin = DiD - 1.96 * SE,
    ymax = DiD + 1.96 * SE,
    color = sex,
    shape = q < 0.05,
    group = sex
  )
) +
  geom_hline(yintercept = 0, linewidth = 0.4, color = "grey40") +
  geom_pointrange(position = position_dodge(width = 0.7), size = 0.15, fatten = 1.4) +
  facet_grid(category ~ timepoint, scales = "free_x", space = "free_x") +
  scale_color_manual(values = c(Female = "#CC79A7", Male = "#0072B2"), name = "Sex") +
  scale_shape_manual(values = c(`FALSE` = 1, `TRUE` = 16), name = "FDR q < 0.05") +
  labs(
    x = "Syllable ID",
    y = "DiD (logit)",
    title = "Sex-specific CORT vs VEH DiD by syllable category",
    subtitle = "Estimate ± 1.96 SE; filled = FDR q < 0.05"
  ) +
  theme_bw(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6),
    strip.background = element_rect(fill = "grey95", color = NA),
    panel.grid.minor = element_blank()
  )

print(p_did)
save_plot_eps_png(p_did, "did_by_category", width = 12, height = 8)
```

```{r did-category-time-plot, eval=FALSE}
# Optional denser alternative kept for exploration; main figure is above.
```

## DDD by category (one plot per category)

```{r ddd-category-plots, fig.width=6.5, fig.height=7}
ddd_theme <- theme_bw(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8),
    strip.background = element_rect(fill = "grey95", color = NA),
    panel.grid.minor = element_blank()
  )

for (cat in cat_levels) {
  ddd_cat <- plot_df %>%
    filter(category == cat) %>%
    mutate(syllable = factor(syllable, levels = sort(unique(syllable))))

  if (nrow(ddd_cat) == 0) {
    cat("No plottable DDD rows for ", cat, "\n", sep = "")
    next
  }

  n_syll <- n_distinct(ddd_cat$syllable)
  fig_w <- max(5.5, 0.35 * n_syll + 2.5)
  fig_h <- 7

  p_ddd <- ggplot(
    ddd_cat,
    aes(
      x = syllable,
      y = DDD,
      ymin = DDD - 1.96 * DDD_SE,
      ymax = DDD + 1.96 * DDD_SE,
      color = DDD_q < 0.05
    )
  ) +
    geom_hline(yintercept = 0, linewidth = 0.5, color = "grey40") +
    geom_pointrange(size = 0.25, fatten = 1.5) +
    facet_grid(timepoint ~ ., switch = "y") +
    scale_color_manual(
      values = c(`FALSE` = "grey60", `TRUE` = "#0072B2"),
      name = "FDR q < 0.05"
    ) +
    labs(
      x = "Syllable ID",
      y = "DDD (logit)",
      title = paste0("DDD: ", str_to_title(cat), " syllables"),
      subtitle = "Male DiD − Female DiD; estimate ± 1.96 SE"
    ) +
    ddd_theme

  print(p_ddd)
  save_plot_eps_png(p_ddd, paste0("ddd_", cat), width = fig_w, height = fig_h)
}
```

## Heatmap summary (one plot per category)

Color = estimate, asterisk = FDR q < 0.05. Shared color scale across categories for comparability.

```{r heatmap-by-category, fig.width=7.5, fig.height=4.5}
heat_df <- plot_df %>%
  select(
    syllable, category, timepoint, DiD_Female, DiD_Male, DDD,
    DiD_Female_q, DiD_Male_q, DDD_q
  ) %>%
  pivot_longer(
    cols = c(DiD_Female, DiD_Male, DDD),
    names_to = "contrast",
    values_to = "estimate"
  ) %>%
  mutate(
    q = case_when(
      contrast == "DiD_Female" ~ DiD_Female_q,
      contrast == "DiD_Male" ~ DiD_Male_q,
      TRUE ~ DDD_q
    ),
    contrast = factor(contrast, levels = c("DiD_Female", "DiD_Male", "DDD")),
    sig = q < 0.05
  )

lim <- max(abs(heat_df$estimate), na.rm = TRUE)

for (cat in cat_levels) {
  heat_cat <- heat_df %>%
    filter(category == cat) %>%
    mutate(syllable = factor(syllable, levels = sort(unique(syllable))))

  if (nrow(heat_cat) == 0) {
    cat("No plottable heatmap rows for ", cat, "\n", sep = "")
    next
  }

  # Height scales lightly with number of syllables in the category
  n_syll <- n_distinct(heat_cat$syllable)
  fig_h <- max(3.2, 0.28 * n_syll + 1.8)

  p_heat <- ggplot(heat_cat, aes(x = timepoint, y = syllable, fill = estimate)) +
    geom_tile(color = "white", linewidth = 0.2) +
    geom_text(aes(label = if_else(sig, "*", "")), size = 3.2, vjust = 0.75) +
    facet_wrap(~contrast, nrow = 1) +
    scale_fill_gradient2(
      low = "#2166AC", mid = "white", high = "#B2182B",
      midpoint = 0, limits = c(-lim, lim), name = "logit"
    ) +
    labs(
      x = "Timepoint",
      y = "Syllable ID",
      title = paste0("DiD / DDD heatmap: ", str_to_title(cat), " syllables"),
      subtitle = "* FDR q < 0.05"
    ) +
    theme_bw(base_size = 11) +
    theme(
      panel.grid = element_blank(),
      strip.background = element_rect(fill = "grey95", color = NA)
    )

  print(p_heat)
  save_plot_eps_png(p_heat, paste0("heatmap_", cat), width = 7.5, height = fig_h)
}
```

## Tables of exact DiD / DDD values

Yes — report the numeric estimates. Main text/figure for pattern; table/supplement for values.

```{r ddd-table}
effects %>%
  filter(estimable, !is.na(category)) %>%
  arrange(DDD_q, DDD_p) %>%
  select(syllable, category, timepoint, DDD, DDD_SE, DDD_z, DDD_p, DDD_q) %>%
  head(30) %>%
  knitr::kable(
    digits = 3,
    caption = "Top DDD effects (male DiD − female DiD), sorted by FDR q"
  )
```

```{r did-tables}
effects %>%
  filter(estimable, !is.na(category), DiD_Female_q < 0.05) %>%
  arrange(category, timepoint, DiD_Female_q) %>%
  select(syllable, category, timepoint, DiD_Female, DiD_Female_SE, DiD_Female_p, DiD_Female_q) %>%
  knitr::kable(digits = 3, caption = "Female DiD with FDR q < 0.05")

effects %>%
  filter(estimable, !is.na(category), DiD_Male_q < 0.05) %>%
  arrange(category, timepoint, DiD_Male_q) %>%
  select(syllable, category, timepoint, DiD_Male, DiD_Male_SE, DiD_Male_p, DiD_Male_q) %>%
  knitr::kable(digits = 3, caption = "Male DiD with FDR q < 0.05")
```

# 6. Interpretation notes

- Estimates are conditional on the mouse random intercept and use a logit link, so they are log-odds contrasts, not raw usage differences.
- $\mathrm{DiD}_{\mathrm{Female},j}$ is the treatment-by-time interaction for females.
- $\mathrm{DDD}_j$ tests whether that longitudinal CORT effect differs in males.
- Sessions with malformed batch `SubjectName` values were remapped to animal IDs from `SessionName` before fitting.
- Rare syllables can yield unstable fits; figures exclude high-SE / extreme estimates, while the CSV retains all fits.
- For a paper: main figure = category-faceted DiD (both sexes) + DDD; supplement = full `did_ddd_effects.csv` and heatmap.
- All plots are written to `figures/eps/` and `figures/png/`.

# 7. Diagnose syllables 15 and 16

These two syllables often have non-finite DiD SEs (`estimable = FALSE`). The block below prints cell-level sparsity, optimizer status, Hessian positive-definiteness, and the full `glmmTMB` summary.

```{r diagnose-15-16}
diagnose_syllable <- function(s) {
  sdat <- did_df %>%
    filter(syllable == s)

  cell_counts <- sdat %>%
    group_by(treatment, sex, timepoint) %>%
    summarise(
      n_sessions = n(),
      total_y = sum(y),
      total_N = sum(N),
      pooled_usage = total_y / total_N,
      zero_sessions = sum(y == 0),
      all_zero = all(y == 0),
      .groups = "drop"
    )

  fit <- glmmTMB(
    cbind(y, N - y) ~
      treatment * timepoint * sex +
      (1 | mouse_id),
    data = sdat,
    family = betabinomial()
  )

  cat("\n===== Syllable ", s, " =====\n", sep = "")
  print(cell_counts)
  cat("\nOptimizer code:", fit$fit$convergence, "\n")
  cat("Optimizer message:", fit$fit$message, "\n")
  cat("Positive-definite Hessian:", fit$sdr$pdHess, "\n\n")

  print(summary(fit))

  if ("diagnose" %in% getNamespaceExports("glmmTMB")) {
    glmmTMB::diagnose(fit)
  }

  invisible(fit)
}

fit15 <- diagnose_syllable(15)
fit16 <- diagnose_syllable(16)
```

# Session info

```{r session-info}
sessionInfo()
```
